### 2018-04-10 latex/icpp18/algorithm.tex
partitioning done

\subsubsection{Partitioning $\Wm$ and $\Hm$}
Once $\Am$ is partitioned with an $R \times C$ checkerboard topology, one has to partition rows and columns of factor matrices conformally with this topology to form the sets $\Ip$ and $\Jp$ in \cref{alg:distnmf}.
In doing so, we are interested in assigning rows and columns to processes in the corresponding process row or column equitably.
For this purpose, we specify imbalance parameters $\alpha_r$ and $\alpha_c$ that correspond to maximum imbalance we allow in this partitioning; i.e., $|\Ip| \le \alpha_r m / P$ and $|\Jp| \le \alpha_c n / P$ for each process $p$.
We normally set these imbalance parameters to $1.1$.
However, when $m >> n$, we increase $\alpha_c$ proportionally in order to give more flexibility in column partitioning which results in smaller communication volume.

Next, for each row and column of $\Wm$ and $\Hm$ we create a list of processes that has a dependency to that row or column, which corresponds to processes owning the matrix blocks of same color in \cref{fig:partition}.
Finally, we randomly assign each row and column to one of the processes satisfying the imbalance constraint in this list.
If all processes in the list are overloaded, we randomly assign it to a process that is not overloaded in the same process row or column.
Note that the latter assignment increases the communication volume due to that row or column by 1; hence, in general larger imbalance parameters yield smaller communication volume by reducing this type of assignment.

>>>>

\subsubsection{Partitioning $\Wm$ and $\Hm$}
Once $\Am$ is partitioned, one has to partition rows and columns of factor matrices to form the sets $\Ip$ and $\Jp$ in \cref{alg:distnmf}.
In doing so, we are interested in assigning rows and columns to processes equitably.
For this purpose, we specify imbalance parameters $\alpha$ that correspond to maximum imbalance we allow in this partitioning; i.e., $|\Ip| \le \alpha m / P$ and $|\Jp| \le \alpha n / P$ for each process $p$, and set $\alpha = 1.05$ in the experiments.

Next, for each row and column of $\Wm$ and $\Hm$ we create a list of processes that has a dependency to that row or column, which corresponds to processes owning the matrix blocks of same color in \cref{fig:partition}.
Finally, we randomly assign each row and column to one of the processes satisfying the imbalance constraint in this list.
If all processes in the list are overloaded, we assign it to the process that has the minimum number of rows/columns assigned.
For a checkerboard partition, the minimum is always chosen from the same processor row/column so that 2D communication topology is not disturbed.
Note that such an assignment increases the communication volume due to that row or column by 1; hence, in general smaller imbalance parameters yield larger communication volume due to increasing this type of assignment.
On the other hand, we desire to keep $\alpha$ small as it pertains to the load imbalance in the NNLS step.
